---
title: "Assignment1"
author: "Paul Weissburg"
date: "May 15, 2019"
output: 
  html_document:
    keep_md: true
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE}
library(dplyr)
library(tidyverse)
library(lubridate)
```

* library(dplyr)
* library(tidyverse)
* library(lubridate)

### 1. Load data into R
Create temp file in WD
```{r}
temp <- tempfile()
```
Download file from URL to temp file

```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
```
Load data into R
```{r}
data <- as.tbl(read.csv(unz(temp, "activity.csv")))
```
Remove temp file
```{r}
unlink(temp)
```
### 2. Histogram of # of steps taken each day
```{r hist}
steps1 <- data %>%
          group_by(date) %>%
          summarise(sum(steps,na.rm = TRUE))
names(steps1) <- c("date","steps")
qplot(steps1$steps,bins=30)
```

### 3. Median and mean of steps taken each day

```{r}
summary(steps1$steps)
```
median = 10395; mean = 9354

### 4. Time series plot of number of steps taken
```{r tSeries, include=TRUE}
dataNArm <- data[!is.na(data$steps),]
ggplot(dataNArm, aes(interval,steps))+geom_point()
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
filter(data, data$steps == max(data$steps,na.rm = TRUE))
```
Interval number 615 has the most steps (806 steps)

### 6. Imputing Missing Values
6a. Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(data$steps))
```
6b. Devise a strategy to impute NA values: Use median for each interval  
6c. Create imputed data sets  

```{r}
dataNew <- data %>%
  group_by(interval) %>%
  mutate(median = median(steps,na.rm = TRUE))%>%
  mutate(steps2 = case_when(is.na(steps) ~ median,!is.na(steps) ~ steps))%>%
  mutate(day = weekdays(as.Date(date)))%>%
  mutate(weekend = case_when(day == "Saturday" | day == "Sunday" ~ "Weekend",
                             TRUE ~ "Weekday"))
dataNew$weekend <- factor(dataNew$weekend,levels = c("Weekday","Weekend"))

dataWeekend <- group_by(dataNew,interval,weekend)%>%
  summarise(mean = mean(steps2))

dataNew2 <- group_by(dataNew,date)%>%
  summarise(sum = sum(steps2))
```
6d. Make a histogram with imputed data
```{r}
qplot(dataNew2$sum,bins = 30)

```

6e. Find median and mean for imputed data
```{r}
summary(dataNew2$sum)

```

median = 10395
mean = 9503.869

6f. The median is the same as the non-imputed calculation but
the mean is higher.

### 7. What is the average daily activity pattern?
Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") 
of the 5-minute interval (x-axis) and the average number of steps taken,
averaged across all days (y-axis)
```{r}
gg <- ggplot(dataWeekend,aes(interval,mean))
gg + geom_line()+facet_grid(weekend~.)
```

Which 5-minute interval, on average across all the days in the dataset, 
contains the maximum number of steps?
```{r}
dataMean <- group_by(dataNew,interval)%>%
  summarise(mean = mean(steps2))
dataMean[dataMean$mean == max(dataMean$mean),]
```
Interval 835 with mean 182 ***Calculated with imputed data***
